---
title: "01_dataset_creation"
author: "Miguel Ángel Armengol"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  # html_notebook:
    code_folding: hide
    number_sections: yes
    theme: flatly
    toc: yes
    toc_float: yes

knit: (function(inputFile, encoding) { rmarkdown::render(inputFile, encoding = encoding, output_file = paste0(substr(inputFile,1,nchar(inputFile)-4)," ",Sys.Date(),'.html')) })
---

# Environment

```{r message=FALSE, warning=FALSE}
library(bigrquery)
library(summarytools)
library(readr)
library(magrittr)
library(sqldf)
library(dplyr)
library(tableone)
library(kableExtra)
library(diagram)
impute.median <- function(x) replace(x, is.na(x), median(x, na.rm = TRUE))
options(scipen=999)
```


# Set up BigQuery related functions

This chunks also creates the run_query and get_sql function.

```{r setup, include=FALSE}
# Updated for our year
project_id <- "hst-953-2018"
options(httr_oauth_cache=FALSE)
# Function that takes in a sql command and runs it on bigquery
run_query <- function(query){
  data <- bq_table_download(bq_project_query(x=project_id,query=query, use_legacy_sql=FALSE,max_pages = Inf))
  return(data)
}

# function for reading sql files
getSQL <- function(filepath){
  con = file(filepath, "r")
  sql.string <- ""

  while (TRUE){
    line <- readLines(con, n = 1)

    if ( length(line) == 0 ){
      break
    }

    line <- gsub("\\t", " ", line)

    if(grepl("--",line) == TRUE){
      line <- paste(sub("--","/*",line),"*/")
    }

    sql.string <- paste(sql.string, line)
  }

  close(con)
  return(sql.string)
}

'%!in%' <- function(x,y)!('%in%'(x,y))
```

# Extracting data

Loads all queries from the sql files in the extraction folder and runs them into RBigQuey to extract the data.

```{r}
apache_aggregated<- run_query(getSQL('sql/apache_aggregated.sql'))
charlson_score<- run_query(getSQL('sql/charlson_score.sql'))
demographics<- run_query(getSQL('sql/demographics.sql'))
lactate_descr <- run_query(getSQL('sql/lactate_descr.sql'))
lactate_first_elev_to_test_mins<-run_query(getSQL('sql/lactate_first_elev_to_test_mins.sql'))
lactate_first_sev_elev_to_test_mins<-run_query(getSQL('sql/lactate_first_sev_elev_to_test_mins.sql'))

lactate_max_first3days<-run_query(getSQL('sql/lactate_max_first3days.sql'))

apache_related<- run_query(getSQL('sql/apache_related.sql'))

# IMPORTANT, ACCORDING TO THE DOCUMMENTATION -1 in Apache score means missing.
apache_related$predictedHospitalMortality[apache_related$predictedHospitalMortality==(-1)] <- NA
apache_related$apache_iv[apache_related$apache_iv==(-1)] <- NA


# SOFA related variables
sofa_cv_day1_to_day4 <- run_query(getSQL("sql/sofa/sofa_cv_day1_to_day4.sql"))
sofa_renal_day1_to_day4 <- run_query(getSQL("sql/sofa/sofa_renal_day1_to_day4.sql"))
sofa_respi_day1_to_day4 <- run_query(getSQL("sql/sofa/sofa_respi_day1_to_day4.sql"))
sofa_3others_day1_to_day4 <- run_query(getSQL("sql/sofa/sofa_3others_day1_to_day4.sql"))
sofa_total_day1_to_day4 <- sqldf((getSQL("sql/sofa/sofa_total_day1_to_day4.sql")))

#qc data
qc_labs<- run_query(getSQL('sql/qc/qc_labs.sql'))
lab_data_qc <- run_query(getSQL('sql/qc/lab_data_qc.sql'))
```

# Joining all datasets

```{r}
# We are using a left join to join them
lactate_df<-Reduce(function(...) merge(..., all.x=TRUE), list(
   demographics
  ,apache_aggregated
  ,apache_related
  ,charlson_score%>%select(patientunitstayid,final_charlson_score)
  ,lactate_descr
  ,lactate_first_elev_to_test_mins
  ,lactate_first_sev_elev_to_test_mins
  ,lactate_max_first3days
  ,sofa_total_day1_to_day4
  ,qc_labs
  ,lab_data_qc
))
```

# Selecting only relevant variables

```{r}
lactate_df
```


# Imputing missing values

```{r}
BMI
body_surface_area
actualicumortality
unabridgedactualventdays
apache_iv
lactate_fst
lactate_max
```


# Exclusion criteria 1

Patients with sepsis (not bleeding) 

```{r}
# we want to include septic patients that are not bleeding.

print('Initial number of patients:')
a <- nrow(lactate_df)
a

print('Patients <18 that are excluded:')
lactate_df %<>% filter(age_fixed > 18)
b <- nrow(lactate_df)
a - b

print('Patients with no Hgb data that are excluded:')
lactate_df %<>% filter(patientunitstayid %in% lab_data_qc$patientunitstayid)
c <- nrow(lactate_df)
b - c

print('Patients with stays shorter than 3 hours that are excluded:')
lactate_df %<>% filter(icu_los_hours >= 3)
d <- nrow(lactate_df)
c - d

print('Patients without a diagnose that are excluded:')
lactate_df %<>% filter(apacheadmissiondx != "")
e <- nrow(lactate_df)
d-e
#####################################################################################################
# we are extracting top n apache groups
topn <- 5
top_apachedxgroup <- tail(names(sort(table(lactate_df$apachedxgroup_mod))), topn) 

lactate_df <- lactate_df %>% mutate(apache_strat = if_else(apachedxgroup_mod %in% top_apachedxgroup, apachedxgroup_mod,'Not Top5'))
#####################################################################################################

print('Patients not in Top5 Apache groups that are excluded:')
lactate_df %<>% filter(apache_strat != "Not Top5")
f <- nrow(lactate_df)
e-f

print('Patients without one lab drawn each day for the first 3 days of admission:')
lactate_df %<>% filter(atleast_one_lab_drawn_each_day_first_3days == 1 )
g <- nrow(lactate_df)
f-g

excluding patients with no mortality

actualicumortality 

print('Final number of patients:')
nrow(lactate_df)
```

## Flowchart diagram

```{r}
if (!requireNamespace("DiagrammeR", quietly = TRUE)) {
  install.packages("DiagrammeR")
}
library(DiagrammeR)

# Define steps and numbers
steps <- c("Initial", "Patients <18", "No Hgb data", "Short stays", "No diagnosis", "Top5 Apache", "Not Top5 Apache", "Lab drawn first 3 days", "Final")
numbers <- c(a, b, c, d, e, f, g, nrow(lactate_df))

# Create a graph
graph <- create_graph(directed = TRUE)

# Add nodes and edges
for (i in 1:(length(steps) - 1)) {
  graph <- add_node(graph, as.character(i), label = steps[i])
  graph <- add_edge(graph, as.character(i), as.character(i + 1), label = as.character(numbers[i] - numbers[i + 1]))
}

graph <- add_node(graph, as.character(length(steps)), label = steps[length(steps)])

# Plot the graph
render_graph(graph)


```

