---
title: "Tables"
author: "Miguel ?ngel Armengol de la Hoz"
output:
  html_document:
    toc: true
    theme: united
---

# Environment

```{r}
library(dplyr)
library(tableone)
library(knitr)
library(kableExtra)
library(flextable)
library(officer)
library(tidyr)

# ============================================================================
# HELPER FUNCTION TO CHECK VARIABLE EXISTENCE
# ============================================================================

check_variables_exist <- function(vars, dataset_name = "lactate_df_selected") {
  existing_vars <- c()
  missing_vars <- c()
  
  for (var in vars) {
    if (var %in% names(lactate_df_selected)) {
      existing_vars <- c(existing_vars, var)
    } else {
      missing_vars <- c(missing_vars, var)
    }
  }
  
  if (length(missing_vars) > 0) {
    cat("Missing variables:", paste(missing_vars, collapse = ", "), "\n")
  }
  
  return(existing_vars)
}

```

# Table 1: Descriptive Data Stratified by Apache Group

```{r}
# ============================================================================
# TABLE 1: DESCRIPTIVE DATA STRATIFIED BY APACHE GROUP
# ============================================================================

create_table1 <- function() {
 
 # Define variables for Table 1 (core descriptive variables)
 vars_in_table1 <- c("age_fixed", "sex", "unabridgedunitlos", "unabridgedhosplos", 
                     "apache_iv", "final_charlson_score", "sofatotal_day1", 
                     "actualicumortality", "region", "teachingstatus")
 
 # Check which variables exist
 vars_in_table1 <- check_variables_exist(vars_in_table1)
 
 # Create Table 1 dataset
 table1_dataset <- lactate_df_selected[, c(vars_in_table1, "apache_strat")]
 stratifyby <- c('apache_strat')
 
 # Set variable labels
 if ("age_fixed" %in% vars_in_table1) attr(table1_dataset$age_fixed, 'label') <- 'Age'
 if ("sex" %in% vars_in_table1) attr(table1_dataset$sex, 'label') <- 'Sex - Male (%)'
 if ("unabridgedunitlos" %in% vars_in_table1) attr(table1_dataset$unabridgedunitlos, 'label') <- 'ICU Length of Stay [Days]'
 if ("unabridgedhosplos" %in% vars_in_table1) attr(table1_dataset$unabridgedhosplos, 'label') <- 'Hospital Length of Stay [Days]'
 if ("apache_iv" %in% vars_in_table1) attr(table1_dataset$apache_iv, 'label') <- 'Apache Score '
 if ("final_charlson_score" %in% vars_in_table1) attr(table1_dataset$final_charlson_score, 'label') <- 'Charlson Comorbidity Index'
 if ("sofatotal_day1" %in% vars_in_table1) attr(table1_dataset$sofatotal_day1, 'label') <- 'SOFA on Day 1'
 if ("actualicumortality" %in% vars_in_table1) attr(table1_dataset$actualicumortality, 'label') <- 'ICU Mortality - n (%)'
 if ("region" %in% vars_in_table1) attr(table1_dataset$region, 'label') <- 'Hospital Region (%)'
 if ("teachingstatus" %in% vars_in_table1) attr(table1_dataset$teachingstatus, 'label') <- 'Academic vs Not (%)'
 
 # Detect categorical variables automatically
 cat_variables <- c()
 for (var in vars_in_table1) {
   if (n_distinct(table1_dataset[[var]], na.rm = TRUE) <= 10) {
     cat_variables <- c(cat_variables, var)
   }
 }
 
 # Create Table 1
 table1_base <- print(CreateTableOne(vars = vars_in_table1,
                                     strata = stratifyby,
                                     includeNA = TRUE,
                                     factorVars = cat_variables,
                                     data = table1_dataset, 
                                     addOverall = TRUE),
                      varLabels = TRUE
                      )
 
 return(table1_base)
}

# Create tables directory if it doesn't exist
if (!dir.exists("./tables")) {
 dir.create("./tables", recursive = TRUE)
}

# Create Table 1
cat("Creating Table 1: Descriptive Data Stratified by Apache Group\n")
table1_result <- create_table1()

# Export Table 1 to Word
export_table_to_word(table1_result, 
                    "./tables/Table1_Descriptive_Apache_Groups.docx",
                    "Table 1: Descriptive Data Stratified by Apache Group")
```

# Table 2: Lactate Data Stratified by Apache Group

```{r}
# ============================================================================
# TABLE 2: LACTATE DATA STRATIFIED BY APACHE GROUP  
# ============================================================================

create_table2 <- function() {
  
  # Define potential variables for Table 2 (lactate-specific variables)
  potential_vars <- c("lactate_bin_first3days", "lactate_num_first3days", 
                      "lactate_num", "lactateredrawn_wt8hrs_bin_first3days",
                      "lactate_fst", "lactate_max_first3days", "lactate_max",
                      "lactate_first", "lactate_maximum", "lactate_count",
                      "lactate_utilization", "lactate_redrawn")
  
  # Check which variables exist and use only those
  vars_in_table2 <- check_variables_exist(potential_vars)
  
  if (length(vars_in_table2) == 0) {
    cat("No lactate variables found. Checking column names...\n")
    lactate_cols <- names(lactate_df_selected)[grep("lactate|lact", names(lactate_df_selected), ignore.case = TRUE)]
    cat("Available lactate-related columns:", paste(lactate_cols, collapse = ", "), "\n")
    return(NULL)
  }
  
  # Create Table 2 dataset
  table2_dataset <- lactate_df_selected[, c(vars_in_table2, "apache_strat")]
  stratifyby <- c('apache_strat')
  
  # Set variable labels (only for existing variables)
  if ("lactate_bin_first3days" %in% vars_in_table2) 
    attr(table2_dataset$lactate_bin_first3days, 'label') <- 'Lactate utilization within first 3 days of hospitalization'
  if ("lactate_num_first3days" %in% vars_in_table2) 
    attr(table2_dataset$lactate_num_first3days, 'label') <- 'Number of lactates drawn (first 3 days)'
  if ("lactate_num" %in% vars_in_table2) 
    attr(table2_dataset$lactate_num, 'label') <- 'Number of lactates drawn (hospitalization)'
  if ("lactateredrawn_wt8hrs_bin_first3days" %in% vars_in_table2) 
    attr(table2_dataset$lactateredrawn_wt8hrs_bin_first3days, 'label') <- 'Number of repeat lactates drawn within 8 hours'
  if ("lactate_fst" %in% vars_in_table2) 
    attr(table2_dataset$lactate_fst, 'label') <- 'Initial Lactate'
  if ("lactate_max_first3days" %in% vars_in_table2) 
    attr(table2_dataset$lactate_max_first3days, 'label') <- 'Maximum Lactate (first 3 days)'
  if ("lactate_max" %in% vars_in_table2) 
    attr(table2_dataset$lactate_max, 'label') <- 'Maximum Lactate (hospitalization)'
  
  # Detect categorical variables automatically
  cat_variables <- c()
  for (var in vars_in_table2) {
    if (n_distinct(table2_dataset[[var]], na.rm = TRUE) <= 10) {
      cat_variables <- c(cat_variables, var)
    }
  }
  
  # Create Table 2
  table2_base <- print(CreateTableOne(vars = vars_in_table2,
                                      strata = stratifyby,
                                      includeNA = TRUE,
                                      factorVars = cat_variables,
                                      data = table2_dataset, 
                                      addOverall = TRUE),
                       varLabels = TRUE
                       )
  
  return(table2_base)
}

# Create Table 2
cat("\nCreating Table 2: Lactate Data Stratified by Apache Group\n")
table2_result <- create_table2()

# Export Table 2 to Word
export_table_to_word(table2_result, 
                     "./tables/Table2_Lactate_Apache_Groups.docx", 
                     "Table 2: Lactate Data Stratified by Apache Group")
```



# Supplemental Table: Additional Variables

```{r}
# ============================================================================
# SUPPLEMENTAL TABLE: REMAINING VARIABLES
# ============================================================================

create_supplemental_table <- function() {
  
  # Define variables for Supplemental Table (remaining variables)
  vars_in_supplemental <- c("unitType", "hospitalAdmitSource", "lactate_max_first3days_type",
                            "hospital_los_hours")
  
  # Check which variables exist
  vars_in_supplemental <- check_variables_exist(vars_in_supplemental)
  
  # Create Supplemental Table dataset
  supplemental_dataset <- lactate_df_selected[, c(vars_in_supplemental, "apache_strat")]
  stratifyby <- c('apache_strat')
  
  # Set variable labels
  if ("unitType" %in% vars_in_supplemental) attr(supplemental_dataset$unitType, 'label') <- 'ICU Type'
  if ("hospitalAdmitSource" %in% vars_in_supplemental) attr(supplemental_dataset$hospitalAdmitSource, 'label') <- 'Hospital Admission Source (%)'
  if ("lactate_max_first3days_type" %in% vars_in_supplemental) attr(supplemental_dataset$lactate_max_first3days_type, 'label') <- 'Maximum Lactate Type in the first Days (%)'
  if ("hospital_los_hours" %in% vars_in_supplemental) attr(supplemental_dataset$hospital_los_hours, 'label') <- 'Hospital Length of Stay (Hours)'
  
  # Detect categorical variables automatically
  cat_variables <- c()
  for (var in vars_in_supplemental) {
    if (n_distinct(supplemental_dataset[[var]], na.rm = TRUE) <= 10) {
      cat_variables <- c(cat_variables, var)
    }
  }
  
  # Create Supplemental Table
  supplemental_base <- print(CreateTableOne(vars = vars_in_supplemental,
                                            strata = stratifyby,
                                            includeNA = TRUE,
                                            factorVars = cat_variables,
                                            data = supplemental_dataset, 
                                            addOverall = TRUE),
                             varLabels = TRUE,
                             nonnormal = c('hospital_los_hours'))
  
  return(supplemental_base)
}

# Create Supplemental Table
cat("\nCreating Supplemental Table: Additional Variables\n")
supplemental_result <- create_supplemental_table()

# Export Supplemental Table to Word
export_table_to_word(supplemental_result, 
                     "./tables/Supplemental_Table_Apache_Groups.docx",
                     "Supplemental Table: Additional Variables Stratified by Apache Group")

cat("\nAll available tables exported to Word documents in ./tables/ directory!")

```
# Table 3: Lactate Utilization by Academic vs Non-Academic Centers Stratified by Disease State

```{r}
# ============================================================================
# TABLE 3: LACTATE UTILIZATION BY ACADEMIC VS NON-ACADEMIC CENTERS
# ============================================================================

create_table3 <- function() {
  
  # Create a dataset with the required variables
  table3_dataset <- lactate_df_selected[, c("lactate_bin_first3days", "apache_strat", "teachingstatus")]
  
  # Convert teachingstatus to meaningful labels
  table3_dataset$academic_status <- ifelse(table3_dataset$teachingstatus == 1, "Academic", "Non-academic")
  
  # Create a combined stratification variable
  table3_dataset$strata_combined <- paste(table3_dataset$apache_strat, table3_dataset$academic_status, sep = " - ")
  
  # Set variable label
  attr(table3_dataset$lactate_bin_first3days, 'label') <- 'Lactate utilization within first 3 days of hospitalization - n (%)'
  
  # Create Table 3 - stratified by apache group AND academic status
  table3_base <- print(CreateTableOne(vars = "lactate_bin_first3days",
                                      strata = c("apache_strat", "academic_status"),
                                      includeNA = FALSE,
                                      factorVars = "lactate_bin_first3days",
                                      data = table3_dataset, 
                                      addOverall = TRUE),
                       varLabels = TRUE)
  
  return(table3_base)
}

# Alternative approach for a more structured cross-tabulation
create_table3_crosstab <- function() {
  
  # Create the dataset
  table3_data <- lactate_df_selected %>%
    mutate(
      academic_status = case_when(
        teachingstatus == 1 ~ "Academic",
        teachingstatus == 0 ~ "Non-academic",
        TRUE ~ "Unknown"
      ),
      lactate_used = case_when(
        lactate_bin_first3days == "1" ~ "Yes",
        lactate_bin_first3days == "0" ~ "No",
        TRUE ~ "Unknown"
      )
    )
  
  # Create cross-tabulation
  crosstab_result <- table3_data %>%
    filter(!is.na(apache_strat) & !is.na(academic_status) & lactate_used != "Unknown") %>%
    group_by(apache_strat, academic_status, lactate_used) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(apache_strat, academic_status) %>%
    mutate(
      total = sum(n),
      percentage = round((n/total) * 100, 1)
    ) %>%
    filter(lactate_used == "Yes") %>%
    select(apache_strat, academic_status, n, total, percentage) %>%
    mutate(result = paste0(n, " (", percentage, "%)"))
  
  # Reshape to wide format
  table3_wide <- crosstab_result %>%
    select(apache_strat, academic_status, result) %>%
    pivot_wider(names_from = academic_status, values_from = result, values_fill = "0 (0.0%)")
  
  return(table3_wide)
}

# Create Table 3
cat("\nCreating Table 3: Lactate Utilization by Academic vs Non-Academic Centers\n")
table3_result <- create_table3()

# Also create the cross-tabulation version
table3_crosstab_result <- create_table3_crosstab()

# Export Table 3 to Word
export_table_to_word(table3_result, 
                     "./tables/Table3_Lactate_Academic_Status.docx",
                     "Table 3: Lactate Utilization by Academic vs Non-Academic Centers Stratified by Disease State")

# Export cross-tabulation version as well
if (!is.null(table3_crosstab_result)) {
  # Convert to format suitable for export_table_to_word
  table3_crosstab_df <- as.data.frame(table3_crosstab_result)
  rownames(table3_crosstab_df) <- table3_crosstab_df$apache_strat
  table3_crosstab_df <- table3_crosstab_df[, -1]  # Remove apache_strat column since it's now rownames
  
  export_table_to_word(table3_crosstab_df, 
                       "./tables/Table3_Lactate_Academic_Crosstab.docx",
                       "Table 3: Lactate Utilization - Academic vs Non-Academic (Cross-tabulation)")
}

cat("Table 3 exported successfully!\n")
```
# Table 4: ICU Mortality Based on Maximum Lactate in First 3 Days Across Different Disease States

```{r}
# ============================================================================
# TABLE 4: ICU MORTALITY BY MAXIMUM LACTATE IN FIRST 3 DAYS
# ============================================================================

create_table4 <- function() {
  
  # Create a dataset with the required variables
  table4_dataset <- lactate_df_selected[, c("actualicumortality", "lactate_max_first3days_type", "apache_strat")]
  
  # Clean the lactate categories if needed
  table4_dataset$lactate_category <- table4_dataset$lactate_max_first3days_type
  
  # Set variable label
  attr(table4_dataset$actualicumortality, 'label') <- 'ICU Mortality - n (%)'
  
  # Create Table 4 - stratified by apache group AND lactate category
  table4_base <- print(CreateTableOne(vars = "actualicumortality",
                                      strata = c("apache_strat", "lactate_category"),
                                      includeNA = FALSE,
                                      factorVars = "actualicumortality",
                                      data = table4_dataset, 
                                      addOverall = TRUE),
                       varLabels = TRUE)
  
  return(table4_base)
}

# Alternative approach for cross-tabulation format
create_table4_crosstab <- function() {
  
  # Create the dataset
  table4_data <- lactate_df_selected
  table4_data$mortality <- ifelse(table4_data$actualicumortality == "1", "Yes", "No")
  
  # Clean lactate categories
  table4_data$lactate_category <- table4_data$lactate_max_first3days_type
  
  # Remove missing data
  table4_clean <- table4_data[!is.na(table4_data$apache_strat) & 
                              !is.na(table4_data$lactate_category) & 
                              !is.na(table4_data$mortality), ]
  
  # Calculate mortality percentages for each combination
  crosstab_result <- table4_clean %>%
    group_by(apache_strat, lactate_category, mortality) %>%
    summarise(n = n(), .groups = "drop") %>%
    group_by(apache_strat, lactate_category) %>%
    mutate(
      total = sum(n),
      percentage = round((n/total) * 100, 1)
    ) %>%
    filter(mortality == "Yes") %>%
    select(apache_strat, lactate_category, n, total, percentage) %>%
    mutate(result = paste0(n, " (", percentage, "%)"))
  
  # Reshape to wide format
  table4_wide <- crosstab_result %>%
    select(apache_strat, lactate_category, result) %>%
    pivot_wider(names_from = lactate_category, values_from = result, values_fill = "0 (0.0%)")
  
  return(table4_wide)
}

# Check what lactate categories we have available
cat("Available lactate categories in lactate_max_first3days_type:\n")
print(table(lactate_df_selected$lactate_max_first3days_type, useNA = "ifany"))

# Create Table 4
cat("\nCreating Table 4: ICU Mortality by Maximum Lactate in First 3 Days\n")
table4_result <- create_table4()

# Also create the cross-tabulation version
table4_crosstab_result <- create_table4_crosstab()

# Export Table 4 to Word
export_table_to_word(table4_result, 
                     "./tables/Table4_Mortality_Lactate_Categories.docx",
                     "Table 4: ICU Mortality Based on Maximum Lactate in First 3 Days Across Different Disease States")

# Export cross-tabulation version as well
if (!is.null(table4_crosstab_result)) {
  # Convert to format suitable for export_table_to_word
  table4_crosstab_df <- as.data.frame(table4_crosstab_result)
  rownames(table4_crosstab_df) <- table4_crosstab_df$apache_strat
  table4_crosstab_df <- table4_crosstab_df[, -1]  # Remove apache_strat column since it's now rownames
  
  export_table_to_word(table4_crosstab_df, 
                       "./tables/Table4_Mortality_Lactate_Crosstab.docx",
                       "Table 4: ICU Mortality by Lactate Categories (Cross-tabulation)")
}

cat("Table 4 exported successfully!\n")
```

# Deprecated

## Stratified by top apache groups

```{r}
vars_in_table1<-c("age_fixed", "sex", "hospital_los_hours", "unitType", "hospitalAdmitSource", "actualicumortality", "unabridgedunitlos", "unabridgedhosplos", "apache_iv", "final_charlson_score", "lactate_fst", "lactate_max", "lactate_num", "lactate_num_first3days", "lactate_bin_first3days", "lactateredrawn_wt8hrs_bin_first3days", "lactate_max_first3days_type", "apache_strat", "hrs_from_first_elev_to_test_dec", "mins_from_first_elev_to_test", "sofa_total_day1")

table1_dataset<-lactate_df_selected[,vars_in_table1]


stratifyby<-'apache_strat'

cat_variables<-rep(NA, length(vars_in_table1))
attr(table1_dataset$age_fixed,'label')<-'Age'
# label(table1_dataset$sex)<-'Sex'
# label(table1_dataset$unitType)<-'ICU Type'
# label(table1_dataset$apache_iv)<-'Apache Score'
# label(table1_dataset$final_charlson_score)<-'Charlson Comorbidity Index'


# detects whether a variable is categorical or not
cont<-1
for (i in 1:length(vars_in_table1) ) {
  if ( n_distinct(table1_dataset[vars_in_table1[i] ])<=10 ) {
    print(i)
    print(vars_in_table1[i])
    print(names(table1_dataset[vars_in_table1[i]]))
    cat_variables[cont]<-names(table1_dataset[vars_in_table1[i]])
    cont<-cont+1
  }
}  

vars_in_table1_nostrat<-c("age_fixed","sex","hospital_los_hours","unitType","hospitalAdmitSource","actualicumortality","unabridgedunitlos","unabridgedhosplos","unabridgedactualventdays","apache_iv","final_charlson_score","lactate_fst","lactate_max","lactate_num","lactate_num_first3days","lactate_bin_first3days","lactateredrawn_wt8hrs_bin_first3days","mins_from_first_elev_to_test","lactate_max_first3days_type")

cat_variables<-cat_variables[!is.na(cat_variables)]
table1_base<-print(CreateTableOne(vars = vars_in_table1_nostrat
                                  , strata = stratifyby
                                  , factorVars = cat_variables
                                  
    ,data = table1_dataset, addOverall=T),varLabels = T
    ,nonnormal = c('leaking_index'))
# run this in console for html output, the code below uses kableExtra::
starification_cats<-n_distinct(table1_dataset[,stratifyby])
table1_base %>%
  kbl(caption = "Table 1" , align = "c") %>%
  kable_classic_2(full_width = F, html_font = "Cambria")%>%
  add_header_above(c("Top apache groups" = 9 ))
```

## Exploratory Table 

### Stratified by top apache groups & lactate_max_first3days_type

```{r}
vars_in_table1<-c("age_fixed", "sex", "hospital_los_hours", "unitType", "hospitalAdmitSource", "actualicumortality", "unabridgedunitlos", "unabridgedhosplos", "apache_iv", "final_charlson_score", "lactate_fst", "lactate_max", "lactate_num", "lactate_num_first3days", "lactate_bin_first3days", "lactateredrawn_wt8hrs_bin_first3days", "lactate_max_first3days_type", "apache_strat", "hrs_from_first_elev_to_test_dec", "sofatotal_day1", "teachingstatus", "region")

table1_dataset<-lactate_df_selected[,vars_in_table1]


stratifyby<-c('lactate_max_first3days_type','apache_strat')

cat_variables<-rep(NA, length(vars_in_table1))
attr(table1_dataset$age_fixed, 'label') <- 'Age'
attr(table1_dataset$sex, 'label') <- 'Sex'
attr(table1_dataset$hospital_los_hours, 'label') <- 'Hospital Length of Stay (Hours)'
attr(table1_dataset$unitType, 'label') <- 'Unit Type'
attr(table1_dataset$hospitalAdmitSource, 'label') <- 'Hospital Admission Source'
attr(table1_dataset$actualicumortality, 'label') <- 'ICU Mortality'
attr(table1_dataset$unabridgedunitlos, 'label') <- 'Unit Length of Stay'
attr(table1_dataset$unabridgedhosplos, 'label') <- 'Hospital Length of Stay'
attr(table1_dataset$apache_iv, 'label') <- 'APACHE IV'
attr(table1_dataset$final_charlson_score, 'label') <- 'Final Charlson Score'
attr(table1_dataset$lactate_fst, 'label') <- 'Lactate First Measurement'
attr(table1_dataset$lactate_max, 'label') <- 'Maximum Lactate'
attr(table1_dataset$lactate_num, 'label') <- 'Number of Lactate Measurements'
attr(table1_dataset$lactate_num_first3days, 'label') <- 'Number of Lactate Measurements in the first 3 Days'
attr(table1_dataset$lactate_bin_first3days, 'label') <- 'Any lactate available in the first 3 Days'
attr(table1_dataset$lactateredrawn_wt8hrs_bin_first3days, 'label') <- 'Any Lactate Redrawn within 8 Hours in the first 3 Days'
attr(table1_dataset$lactate_max_first3days_type, 'label') <- 'Maximum Lactate Type in the first Days'
attr(table1_dataset$apache_strat, 'label') <- 'APACHE Stratification'
attr(table1_dataset$hrs_from_first_elev_to_test_dec, 'label') <- 'Decile of Hours from First Elevated Lactate to Test'
attr(table1_dataset$sofatotal_day1, 'label') <- 'SOFA on Day 1'
attr(table1_dataset$teachingstatus, 'label') <- 'Hospital Teaching Status'
attr(table1_dataset$region, 'label') <- 'Hospital Region'


# detects whether a variable is categorical or not
cont<-1
for (i in 1:length(vars_in_table1) ) {
  if ( n_distinct(table1_dataset[vars_in_table1[i] ])<=10 ) {
    print(i)
    print(vars_in_table1[i])
    print(names(table1_dataset[vars_in_table1[i]]))
    cat_variables[cont]<-names(table1_dataset[vars_in_table1[i]])
    cont<-cont+1
  }
}  

vars_in_table1_nostrat<-c("age_fixed", "sex", "hospital_los_hours", "unitType", "hospitalAdmitSource", "actualicumortality", "unabridgedunitlos", "unabridgedhosplos", "apache_iv", "final_charlson_score", "lactate_fst", "lactate_max", "lactate_num", "lactate_num_first3days", "lactate_bin_first3days", "lactateredrawn_wt8hrs_bin_first3days", "lactate_max_first3days_type", "apache_strat", "hrs_from_first_elev_to_test_dec", "sofatotal_day1", "teachingstatus", "region")

cat_variables<-cat_variables[!is.na(cat_variables)]
table1_base<-print(CreateTableOne(vars = vars_in_table1_nostrat
                                  , strata = stratifyby
                                  , includeNA = TRUE
                                  , factorVars = cat_variables
                                  
    ,data = table1_dataset, addOverall=T),varLabels = T 
    ,nonnormal = c('leaking_index'))
# run this in console for html output, the code below uses kableExtra::
starification_cats<-n_distinct(table1_dataset[,stratifyby])

table1_base %>%
  kbl(caption = "Table 1" , align = "c") %>%
  kable_classic_2(full_width = F, html_font = "Cambria")
  # add_header_above(c("Top apache groups" = 7 ))


```

### Including missing values

```{r}

vars_to_include_stratified <- ~ age_fixed + sex + hospital_los_hours + unitType + 
  hospitalAdmitSource + actualicumortality + unabridgedunitlos + 
  unabridgedhosplos + apache_iv + final_charlson_score + lactate_fst + 
  lactate_max + lactate_num + lactate_num_first3days + 
  lactate_bin_first3days + lactateredrawn_wt8hrs_bin_first3days + 
  hrs_from_first_elev_to_test_dec + sofatotal_day1 + teachingstatus | 
  lactate_max_first3days_type + apache_strat

# Render the table with stratification
render.table1 <- table1(vars_to_include_stratified, data = lactate_df_selected_withmissinig
, 
                        factorVars = cat_variables)

# Example of direct rendering in R Markdown:
render.table1
```


